Molecular dynamics simulation of synchronization in driven particles

### 2020-10-22 manuscript/manuscript_ajp_2020.tex
commented/outline

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  With
  numerical simulations, we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  We use molecular dynamics simulations to model the particle dynamics,
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We study 
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%

%Outline
%describe types of physical systems 
Numerical simulations of confined, driven particles
can be used to model a diverse variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as wires and fluid microchannels.
Biological systems such as
neuron axons and capillaries can also be studied
with these models [more detail and references].
Many such systems execute local oscillations
about stable points [elaborate].
%
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

Colloidal particles trapped in light fields have proven
a particularly useful medium for studying these behaviors.%cite 
The relatively large size of the colloids
and ease in control has lead to a rich array of 
experimental results. %THIS IS A TERRIBLE SENTENCE.
Such studies are considered model systems
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

Synchronization has been studied for over three centuries,
and is observed in different forms
from Huygens pendulum clocks
to the rhythmic beats of the flapping wings in a flock of birds.
Controlling
synchronization phenomena in weakly coupled oscillators
can be achieved with an external driving force
that causes the syncing of natural oscillation frequencies,
dynamic phase locking \cite{juniper2015}.
[useful and interesting because?]

Disordered chaotic dynamics are also possible,
where irregular, unpredictable time evolution of
nonlinear systems and occurs in mechanical oscillators \cite{chaos}.

In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD},
including various particle interactions and confining environments.
In Section~\ref{sec:sync} we
demonstrate how uniform environments and applied forces
and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams.
We modify the environment and applied forces
and show aperiodic, or nearly periodic flow behaviors
in Section~\ref{sec:quasiperiod}.
Finally we explore the transition to chaos in 
in Section~\ref{sec:chaos}.

@@

\section{Molecular Dynamics Model}
\label{sec:MD}
We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles. 
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5$.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$.
We use periodic boundary conditions
such that a particle leaving the edges of the system is mapped
back to a position within the simulation 
by the rules $x_i+L=x_i$ and $y_i+L=y_i$.

>>>>

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  Using 
  numerical simulations
  we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  The molecular dynamics simulations
  model the particle dynamics 
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We observe
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

\section{Introduction} %
%
%Outline
%-----------------------------------------------------------
%describe types of physical systems - particle types
Numerical simulations of confined, driven particles
can be used to model a variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

%-----------------------------------------------
%environments
Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as wires and fluid microchannels.
Biological systems such as
neuron axons and capillaries can also be studied
with these models [more detail and references].
Many such systems execute local oscillations
about stable points [elaborate].
%----------------------------------------------------------
%external forces
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

%-------------------------------------------------------------
%experimental - colloids specifically
Colloidal particles trapped in light fields have proven
a particularly useful medium for studying these behaviors.%cite 
The relatively large size of the colloids
and ease in control has lead to a rich array of 
experimental results. %THIS IS A TERRIBLE SENTENCE.
Such studies are considered model systems
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

%------------------------------------------------------------------
%synchronization
Synchronization has been studied for over three centuries,
and is observed in different forms
from Huygens pendulum clocks
to the rhythmic beats of the flapping wings in a flock of birds.
Controlling
synchronization phenomena in weakly coupled oscillators
can be achieved with an external driving force
that causes the syncing of natural oscillation frequencies,
dynamic phase locking \cite{juniper2015}.
[useful and interesting because?]

%---------------------------------------------------------------
%kinks and nanofriction

%-----------------------------------------------------------------
%transition from synchronized to chaotic
Disordered chaotic dynamics are also possible,
where irregular, unpredictable time evolution of
nonlinear systems and occurs in mechanical oscillators \cite{chaos}.

%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD},
including various particle interactions and confining environments.
In Section~\ref{sec:sync} we
demonstrate how uniform environments and applied forces
and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of particle velocity vs. position.
We modify the environment and applied forces
and show aperiodic, or nearly periodic flow behaviors
in Section~\ref{sec:quasiperiod}.
Finally we explore the transition to chaos in 
in Section~\ref{sec:chaos}.

@@

\section{Molecular Dynamics Model}
\label{sec:MD}
We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles. 
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5$.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$.
Particles are subject to
periodic boundary conditions
such that a particle leaving the edges of the system is mapped
back to a position within the simulation 
by the rules $x_i+L=x_i$ and $y_i+L=y_i$.
